---
title: "**EDA - Solar Energy Prediction**"
author: "GROUP F: ADDISON P., ALBERTO I., DANIEL R., FIDELIS M., VANIA C."
date: "May 22nd 2020"
output:
  html_document:
    fig_caption: yes
    toc: yes
    toc_float: yes
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    toc: yes
---


```{r, warning=FALSE, echo=FALSE, results='hide'}
library(rmarkdown)
library(knitr)
library(data.table)
library(ggplot2)
library(leaflet)
library(maps)
library(Amelia)
opts_chunk$set(warning=FALSE, echo=FALSE, fig.width = 15)
```


```{r,results='hide'}

########################## GLOBAL PATHS ##########################
folder_path <- "../files/";

############################# GLOBAL QUERYS #############################
data <- readRDS(file.path(folder_path, "solar_dataset.RData"));
stations <- fread(file.path(folder_path, "station_info.csv"));
additional_vars <- readRDS(file.path(folder_path, "additional_variables.RData"));

### Station names
stationsNames <- colnames(data)[2:99]
data <- cbind(data, month = as.numeric(sapply(data$Date, substr, 5,6)));
data <- data[1:5113,]

#Select Principal Components
pc <- colnames(data)[100:456];

```

## Solar dataset

The dimension of Solar dataset is `r nrow(data)` x `r ncol(data)`. 

First column is `r colnames(data)[1]` from `r as.Date(data$Date[1], format = "%Y%m%d")` to `r as.Date(data$Date[5113], format = "%Y%m%d")`. 

Next `r length(stationsNames)` contain energy production in solar estations in units of Joules per square meter. 

Finally, there are `r length(pc)` columns with Principal Components that can be used as predictors.



### Location of the 98 Solar Stations


The following map shows the location of the stations. 

```{r,results="asis"}  

mapStates = map("state", fill = TRUE, plot = FALSE)

m <- leaflet(data = mapStates, width = 800, height = 400) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  setView(-97.34, 35.54, zoom = 10) %>%
  fitBounds(-104.00, 38.80, -94.00, 33.20) %>%
  addPolygons(fillColor = topo.colors(10, alpha = NULL), stroke = FALSE) %>%
  addCircles(data = stations, lng = ~elon, lat = ~nlat, popup = ~stid, color = 'red');
m 
```

Click the red dots to know the name of a particular station. 


### Solar Stations Data  

Number of missing values: 

```{r}
sum(is.na(data));
```


As the stations are close between them, their data are very similar. 

```{r}
boxplot(data[, ..stationsNames], main = "Boxplot for the 98 stations") #boxplot 
```

There is a station between HOOK and LAHO that seems to have an unexpected high value, let´s see
which it is:

```{r}
boxplot(data[, HOOK:LAHO], main = "Boxplot from HOOK to LAHO") #boxplot 
```

ACME and IDAB are used as examples to analize stations data:

Summary

```{r}
summary(data[, .(ACME, IDAB)])
```

Density function to see how often a particular energy value appears:

```{r}
plot(density(data$ACME), col = "blue", main = "Density function ACME")
plot(density(data$IDAB), col = "red", main = "Density function IDAB")

ggplot() + 
geom_density(aes(x = data[,ACME]), colour = "blue") +
geom_density(aes(x= data[, IDAB]), colour = "red") +
labs(x = "Energy (joules per squared meter)", y = "Frecuency") +
theme_bw()  
  
```

If we plot all the data from a Solar Station we can see the seasonality of energy production: 

```{r}
ggplot(data) +
geom_line(aes(x = Date, y = ACME, group = 1), color = "orange") +
theme_bw() +
theme(axis.text.x = element_blank()) +
labs(x = "Years 1994 to 2007", y = "Solar Station Production")
``` 

If we zoom in a year we can see the following pattern: 

```{r}
ggplot(data[Date %in% "19940101":"19941231"]) +
  geom_point(aes(x = month, y = ACME, color = month)) +
  theme_bw() +
  labs(x = "Month of the year - 1994", y = "Solar Station Production")+
  scale_x_discrete(labels = 1:12, limits = c(1:12))+
  scale_color_gradient2(midpoint=6, low="blue", mid="orange",
                        high="blue")
```
The mean of the energy production in the month of May from 1994 to 2007 has been: 
`r formatC(mean(as.matrix(data[month == 5, ACME:WYNO])), format = "e", digits = 2)` while in the month of December it has been: 
`r formatC(mean(as.matrix(data[month == 12, ACME:WYNO])), format = "e", digits = 2)`.

### Principal Components Analysis

```{r,results="asis"}

variance <- sapply(data[, ..pc], var);
prop_var <- variance / sum(variance);
prop_var_acum <- cumsum(prop_var);

ggplot() +
  geom_col(aes(x = 1:10 , y = prop_var[1:10]), width = 0.7) +
  geom_line(aes(x = 1:10 , y = prop_var_acum[1:10]), col = "blue") +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  labs(x = "Principal Component", y = "Percentage of Explained Variance") +
  scale_x_discrete(labels = paste0("PC", seq(1,10)), limits = c(1:10))

```
80% of the variance is explained by the first `r which(prop_var_acum >= 0.8)[1]` Principal Components

### Additional Variables Analysis

Total missing values in Additional variables dataset are `r sum(is.na(additional_vars[, -"Date"]))`

```{r}
max(sapply(additional_vars[,-"Date"], function(x){100 * (sum(is.na(x))) / length(x)}));
```

Up to 7% NAs exist for additional_vars (additional_variables dataset)

```{r}
missmap(additional_vars);
```

Additional variables have skwed distributions with more that 75% of values = 0 and many sparse outliers:
```{r}
plot(density(additional_vars$V649, na.rm = TRUE), col = "blue", main = "Density function example")
```
This is why the team decided not to use any of these additional variables for the Machine Learning model.